---
title: "Literate Programming"
output: html_document
date: "2023-09-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Include Tidyverse library
library(tidyverse)

# Import Data

indata <- read.csv("C:/Users/somes/Downloads/FertilityRates.csv")

# Check if data import was as expected
dim(indata)
head(indata)

# Is it a data frame or a tibble?
is.data.frame(indata)

indata <- as_tibble(indata)
is_tibble(indata)

# Summary of data
summary(indata)
```

The data shows us the fertility rate around the world from 1960 to 2011, it has a total of 219 rows (120 countries).

```{r}
# Convert Country Name into a factor
indata$Country.Name <- as.factor(indata$Country.Name)

summary(indata$Country.Name)

# Convert Indicator Name into a factor
indata$Indicator.Name <- as.factor(indata$Indicator.Name)
summary(indata$Indicator.Name)

# Remove extra columns
indata_cleaned <- select(indata, -(Country.Code:Indicator.Code))

head(indata_cleaned)
```

Storing country names and indicator name as a factor and then deleting it from the loaded data to clean the data



```{r}

# Pivot to long dataset
indata_pivoted <- pivot_longer(indata_cleaned, c(str_c("X", c(1960:2011))), names_to = "Year", values_to="Fertility.Rates")

head(indata_pivoted)

# Format Year values
indata_pivoted$Year <- as.integer(str_sub(indata_pivoted$Year, 2,5))

# Check Missing Values
sum(is.na(indata_pivoted$Fertility.Rates))

# Check missing values by country

## Step1: Filter by missing values
## setp 2: group by country
## Step3: Count number of values

indata_pivoted %>%  filter(is.na(Fertility.Rates)) %>% group_by(Country.Name) %>% summarise(count = n())
```

First we pivot the data from a long format data to wide format data, then we will be checking for any missing values in the data.

```{r}

# Fill missing values within countries

indata_filled <- indata_pivoted %>% group_by(Country.Name) %>% fill(Fertility.Rates, .direction ="downup") %>% ungroup()

## Check
indata_filled %>%  filter(is.na(Fertility.Rates)) %>% group_by(Country.Name) %>% summarise(count = n())

indata_filled <- indata_filled %>% filter(!is.na(Fertility.Rates))

ggplot(data = indata_filled) + geom_point(mapping = aes(x=Year, y = Fertility.Rates))

# Plot Fertility Rates by Year
ggplot(data = indata_filled) + geom_point(mapping = aes(x=Year, y = Fertility.Rates), position = "jitter", alpha = 0.1) + geom_smooth(mapping = aes(x=Year, y = Fertility.Rates)) + labs(x = "Year", y = "Fertility Rates", title = "Fertility Rates over the years", subtitle = "Global fertility rates have decreased since 1960")
```

We are filling in the missing values and then plotting the graph to check if any discrepancies.
After that we plotting the data by the year, overall the fertility rate is decreasing year by year.

```{r}
# Need a way to categorize 210 countries

## Option 1: subset the data

indata_subset <- filter(indata_filled, Country.Name %in% c("United States", "Mexico", "Canada"))


ggplot(data = indata_subset) + geom_line(mapping = aes(x=Year, y = Fertility.Rates, color = Country.Name), alpha = 0.5) + labs(x = "Year", y = "Fertility Rates", title = "Fertility Rates over the years", subtitle = "Global fertility rates have decreased since 1960")
```

plotting subset of the data for the countries "United States", "Mexico", "Canada", its a line graph

```{r}

## Option 2: Select based on statistics

# Top 10

indata_filled %>%  group_by(Country.Name) %>% summarise(avg = mean(Fertility.Rates)) %>% arrange(desc(avg)) %>% print(n = 10)


# Bottom 10

indata_filled %>%  group_by(Country.Name) %>% summarise(avg = mean(Fertility.Rates)) %>% arrange(desc(avg)) %>% top_n(10) %>% ggplot() + geom_bar(mapping = aes(x = Country.Name, y = avg), stat = "identity") +coord_flip()
```

listing top 10 countries and plotting a bar graph for bottom 10 countries

```{r}

### Option
library(countrycode)

indata_df <- as.data.frame(indata_filled)
indata_df$Continent.Name <- factor(countrycode(sourcevar = indata_df[,"Country.Name"], origin = "country.name", destination = "continent"))


indata_df$Region.Name <- factor(countrycode(sourcevar = indata_df[,"Country.Name"], origin = "country.name", destination = "region"))

head(indata_df)

indata <- as_tibble(indata_df)


ggplot(data = indata) + geom_point(mapping = aes(x=Year, y = Fertility.Rates, color = Continent.Name), position = "jitter", alpha = 0.6) + labs(x = "Year", y = "Fertility Rates", title = "Fertility Rates over the years", subtitle = "Global fertility rates have decreased since 1960")

# Facet Plotting
ggplot(data =  indata) + geom_point(mapping = aes(x=Year, y = Fertility.Rates, color = Country.Name), position = "jitter", alpha = 0.2, show.legend = FALSE) + facet_wrap(~ Region.Name)

```

Getting continents for the countries, then plotting them in the same graph based on the continent. Then plotting the graphs individually based on the region.

```{r}

# Filtering and Potting
indata %>% filter(Region.Name == "Middle East & North Africa" & Continent.Name == "Asia") %>% ggplot() + geom_line(mapping = aes(x =Year, y = Fertility.Rates, color =  Country.Name), size = 1, linetype = 2)

# Boxplot

ggplot(data =  indata, mapping = aes(x = Continent.Name, y = Fertility.Rates)) + geom_boxplot() + coord_flip()

# Histogram
ggplot(data = indata) + geom_histogram(mapping = aes(x = Fertility.Rates), binwidth = 0.5)


# Barchart

indata %>% group_by(Region.Name) %>% summarise(avg = mean(Fertility.Rates)) %>% ggplot(mapping = aes(x = reorder(Region.Name, -avg), y = avg, fill=Region.Name)) + geom_bar(stat = "identity") + coord_flip()
```

Plotting the graph for just middle eastern and north african countries.

Plotting box plots for the continents

plotting histogram for all data

the last graph plots the average fertility rate for different regions.